---
title: "MSc_Thesis"
author: "Shannon McNee"
date: "28 November 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(viridis)
library(psych)
library(lavaan)
library(Hmisc)
library(corrplot)
```


## Overview 

In the current study, we are interested in understanding whether an individual's social comparison orientation and social media engagement style relate to levels of night-time social media use, rumination, emotional investment in social media, and sleep outcomes. In this study, social comparison orientation is discussed in relation to two subtypes: ability-based comparison and opinion-based comparison. Social media engagement style is also broken down into two types: active and passive. 

We predict: 

H1. Individuals who engage in passive night-time social media use will demonstrate increased ability-based social comparison. 

H2. Individuals who engage in ability-based social comparison will experience increased emotional investment in social media. 

H3. Individuals who engage in ability-based social comparison will also experience increased rumination, mediated by increases in emotional investment in social media.

H4. Increased rumination will be associated with poor sleep for individuals who engage in ability-based social comparison

Ultimately, we propose a model which predicts that passive night-time social media use will be associated with increased engagement in ability-based comparison (H1). In turn, increased ability-based comparison will be associated with increased emotional investment in social media (H2) which will act as a mediator of the relationship between ability-based comparison and rumination (H3), with increased rumination resulting in reduced sleep outcomes (H4).

## Load in Data 

```{r load dat}

data <- read_csv("Social-media-mental-wellbeing-and-sleep-quests_2019-01-29.csv")

```

## Cleaning and Organising Data

1. Filter data: Remove any observations which are not from registered participants (i.e. admins or students), remove any questionnaires which are not demographic and remove anyone that reports a sleep condition.

```{r clean}

filtered_demographic_data <- data %>%
  select(user_id, user_sex, user_status, user_age, quest_id, q_name,dv) %>%
  filter(user_status == "registered", quest_id == 18, q_name == "fb_use" | q_name == "insta_use" | q_name == "twitter_use" | q_name == "combined_use" | q_name == "sleep_condition" | q_name == "sleep_treatment") %>%
  distinct(user_id,q_name, .keep_all = TRUE) %>%
  spread(q_name,dv) %>%
  filter(combined_use < 7) %>%
  filter(str_detect(sleep_condition,"^N/") | str_detect(sleep_condition,"^n/") | str_detect(sleep_condition, "^N") | str_detect(sleep_condition, "^n")) %>%
gather("q_name","dv", 6:11)

```

2. Filter out demographic questionnaires and rescore questions whose formats were not supported by Experimentum platform. 

```{r recode measures}

q_data <- data %>%
  select(user_id, user_status, quest_id, q_name, dv) %>%
  arrange(user_id) %>%
  filter(user_status == "registered", quest_id < 15, user_id != 76) %>%
  select(user_id, q_name, dv) %>%
  distinct(user_id,q_name, .keep_all = TRUE) %>%
  spread(q_name,dv) %>%
  mutate(SC_Q5 = ifelse(SC_Q5 == 5, 1,
                        ifelse(SC_Q5 == 4,2,
                               ifelse(SC_Q5 == 3, 3,
                                      ifelse(SC_Q5 == 2, 4,
                                             ifelse(SC_Q5 == 1, 5, "")))))) %>%
  mutate(SM2_Q5 = ifelse(SM2_Q5 == 0,1,
                         ifelse(SM2_Q5 == 1,1,
                                ifelse(SM2_Q5 == 2, 0,
                                       ifelse(SM2_Q5 == 3, 0, "")))))

rum <- q_data %>%
  select(R_Q1:R_Q9) %>%
  mutate_all(funs(ifelse(. == 1,0,
                         ifelse(. == 2, 1,
                         ifelse(. == 3,2,
                         ifelse(. == 4, 3,
                         ifelse(. == 5, 4, ""))))))) %>%
  select(R_Q1, R_Q2, R_Q3, R_Q4, R_Q5, R_Q6, R_Q7, R_Q8, R_Q9, R_Q10,
         R_Q11, R_Q12, R_Q13, R_Q14, R_Q15)
 

q_dat <- q_data %>%
  select(-c(R_Q1:R_Q9)) %>%
  cbind(rum) %>%
  gather(q_name,dv,EI_Q1:R_Q15) %>%
  arrange(user_id) %>%
  distinct(user_id,q_name,dv)
  

```

3. Make questionnaire names meaningful   

```{r recode measures2}
questionnaire_data <- q_dat %>%
  mutate(
    questionnaire = ifelse(
      q_name == "SC_Q1" |
        q_name == "SC_Q2" |
        q_name == "SC_Q3" |
        q_name == "SC_Q4" | q_name == "SC_Q5",
      "Ability_Comparison",
      ifelse(
        q_name == "SC_Q6" |
          q_name == "SC_Q7" |
          q_name == "SC_Q8" | q_name == "SC_Q9",
        "Opinion_Comparison",
        ifelse(
          q_name == "PA_Q1" |
            q_name == "PA_Q2" |
            q_name == "PA_Q3" |
            q_name == "PA_Q6" | q_name == "PA_Q12",
          "Active_Social",
          ifelse(
            q_name == "PA_Q5" |
              q_name == "PA_Q7" |
              q_name == "PA_Q9" | q_name == "PA_Q10",
            "Active_NonSoc",
            ifelse(
              q_name == "PA_Q4" |
                q_name == "PA_Q8" |
                q_name == "PA_Q11" | q_name == "PA_Q13",
              "Passive",
              ifelse(
                grepl("R_Q", q_name),
                "Rumination",
                ifelse(
                  grepl("EI_Q", q_name),
                  "Emo_Invest",
                  ifelse(
                    grepl("SM2_Q", q_name),
                    "Night_SM",
                    ifelse(grepl("SCI_Q", q_name), "Sleep_Condition", "")
                  )
                )
              )
            )
          )
        )
      )
    )
  )



questionnaire_names <- unique(questionnaire_data$questionnaire)


```

4. Remove participants where more than half of their responses are incomplete

```{r remove nas}
filtered_questionnaire_data <- questionnaire_data %>%
  group_by(user_id) %>%
  filter(!is.na(dv) > 0.5) %>%
  filter(all(questionnaire_names %in% questionnaire))
  
```

5. Find sum score on each measure per participant and mean, sd, and range of responses 

```{r sums}

sum_scores <- filtered_questionnaire_data %>%
  group_by(user_id, questionnaire) %>%
  summarise(Score = sum(as.integer(dv))) %>%
  arrange(questionnaire) %>%
  ungroup() 

mean_sd_range <- sum_scores %>%
  group_by(questionnaire) %>%
  summarise(Mean = mean(Score),
            SD = sd(Score),
            Min = min(Score),
            Max = max(Score))


```


6. Use this cleaned data to find out age and gender demographic info 

```{r pinfo} 
participant_info <- filtered_demographic_data %>%
  right_join(., sum_scores, by = "user_id") %>%
  select(user_id,user_sex,user_age) %>%
  unique() %>%
  group_by(user_sex) %>%
  summarise(n = n(),
            mean_age = mean(user_age),
            sd_age = sd(user_age))

```

## Confirmatory Analysis

Assess hypothesised relationships between Night-time Social Media Use,Passive Use, Ability Comparison, Emotional Investment, Rumination and Sleep Condition.

### Correlations

```{r corr confirm}

scores <- sum_scores %>%
  spread(questionnaire,Score) %>%
  select(-user_id) 

corr_matrix <- scores %>%
  select(Night_SM,Ability_Comparison, Passive, Emo_Invest, Rumination, Sleep_Condition) %>%
  corr.test(adjust = "none")

```

```{r vis1, echo = TRUE}
# extracting r values from corr_matrix
r_values <- corr_matrix[["r"]] %>%
  as.data.frame() %>%
  rownames_to_column(var = "V1") %>%
  gather("V2","r", Night_SM:Sleep_Condition) 

# extracting p values from corr_matrix

p_values <- corr_matrix[["p"]] %>%
  as.data.frame() %>%
  rownames_to_column(var = "V1") %>%
  gather("V2", "p", Night_SM:Sleep_Condition)

ggplot(r_values, aes(V1,V2, fill = r)) +
         geom_tile(aes(fill = r)) +
         scale_fill_viridis() +
         geom_text(aes(label = paste(round(r,1)))) +
         theme(axis.text.x= element_text(angle=90))

ggplot(p_values, aes(V1,V2, fill = p)) +
         geom_tile(aes(fill = p)) +
         scale_fill_viridis() +
         geom_text(aes(label = paste(round(p,3)))) +
         theme(axis.text.x= element_text(angle=90))


# publication figure 

confirm_dat <- scores %>%
select(Night_SM,Ability_Comparison, Passive, Emo_Invest, Rumination, Sleep_Condition) %>%
cor()

corrplot(confirm_dat, method = "ellipse", type = "upper", diag = FALSE, addCoef.col = "black", tl.col = "black")

```

### Regression

We hypothesised that increased Emotional Investment will act as a mediator between Ability comparison and Rumination 

```{r reg_emo_invest} 


reg3 <- 'Emo_Invest ~ a*Ability_Comparison
         Rumination ~ b*Emo_Invest
         Rumination ~ c*Ability_Comparison

         indirect := a*b
         direct := c
         total := c + (a*b)'

fit_reg3 <- sem(reg3, scores)
summary(fit_reg3, standardized = T, fit.measures = T, rsq = T)


```

### Path Analysis 

We hypothesise that increased levels of passive night-time social media use will be associated with increased levels of ability based comparison. In turn, this will increase emotional investment in social media which will act as a mediator between ability-based comparison and rumination. Increased rumination will then negatively impact sleep. To test this model, we will conduct path analysis. 

```{r path_confirm} 
## Passive Ability

med_mod1 <- 'Ability_Comparison ~ a*Passive
             Emo_Invest ~ b*Ability_Comparison
             Rumination ~ c*Emo_Invest
             Sleep_Condition ~ d*Rumination
             
             
         indirect := b*c
         total := a + (b*c) + d'

fit1 <- sem(med_mod1, scores)
summary(fit1, standardized = T, fit.measures = T, rsq = T)

``` 

## Exploratory Analysis 

As Emotional Investment was not found to be a significant mediator of the relationship between Ability-based Comparison and Rumination, we updated our proposed model and investigated whether Rumination is a significant mediator of the relationship between Ability Comparison and Sleep Outcomes. 

```{r path explore1}
## Passive Ability

med_mod2 <- 'Ability_Comparison ~ a*Passive
             Rumination ~ b*Ability_Comparison
             Sleep_Condition ~ c*Rumination
        

         indirect := b*c
         total := a + (b*c)'

fit2 <- sem(med_mod2, scores)
summary(fit2, standardized = T, fit.measures = T, rsq = T)

```

Passive Use was not found to explain Sleep Outcomes. As a result, we chose to explore whether Active Use (Active-Social versus Active Non-Social) may better explain the relationship between Ability-based comparison, Rumination and Sleep Outcomes. 

### Exploratory Correlations 

```{r corr explore} 

corr_matrix_exp <- scores %>%
  select(Night_SM,Ability_Comparison, Opinion_Comparison, Active_Social, Active_NonSoc, Passive, Emo_Invest, Rumination, Sleep_Condition) %>%
  corr.test(adjust = "none")

```

```{r vis2, echo = TRUE}

# extracting r values from corr_matrix
r_values <- corr_matrix_exp[["r"]] %>%
  as.data.frame() %>%
  rownames_to_column(var = "V1") %>%
  gather("V2","r", Night_SM:Sleep_Condition) 

# visualising r values
ggplot(r_values, aes(V1,V2, fill = r)) +
         geom_tile(aes(fill = r)) +
         scale_fill_viridis() +
         geom_text(aes(label = paste(round(r,1)))) +
         theme(axis.text.x= element_text(angle=90))

# extracting p values from corr_matrix

p_values <- corr_matrix_exp[["p"]] %>%
  as.data.frame() %>%
  rownames_to_column(var = "V1") %>%
  gather("V2", "p", Night_SM:Sleep_Condition)

# visualising p values

ggplot(p_values, aes(V1,V2, fill = p)) +
         geom_tile(aes(fill = p)) +
         scale_fill_viridis() +
         geom_text(aes(label = paste(round(p,3)))) +
         theme(axis.text.x= element_text(angle=90))

# publication figure 

exploratory_dat <- scores %>%
select(Night_SM,Ability_Comparison, Opinion_Comparison, Passive, Active_NonSoc, Active_Social,  Rumination, Sleep_Condition) %>%
cor()

corrplot(exploratory_dat, method = "ellipse", type = "upper", diag = FALSE, addCoef.col = "black", tl.col = "black")


```

### Exploratory Path Analysis

```{r path explore2}

# Active Social Ability

med_mod3 <- 'Ability_Comparison ~ a*Active_Social 
             Rumination ~ b*Ability_Comparison
             Sleep_Condition ~ c*Rumination

             indirect := b*c
             total := a + (b*c)'

fit3 <- sem(med_mod3, scores)
summary(fit3, standardized = T, fit.measures = T, rsq = T)

# Active Non Social Ability

med_mod4 <- 'Ability_Comparison ~ a*Active_NonSoc
             Rumination ~ b*Ability_Comparison
             Sleep_Condition ~ c*Rumination

             indirect := b*c
             total := a + (b*c)'

fit4 <- sem(med_mod4, scores)
summary(fit4, standardized = T, fit.measures = T, rsq = T)

```

